A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-04-09 manuscript.Rmd
added Turing Way citation to Introduction

affiliation:
  - id            : "1"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "2"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany & London, U.K." 
authornote: |
   We are grateful to Julia Delius for her helpful assistance in language and style editing. This paper is fully reproducible using the workflow described in this paper. All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/). This particular document was created from the commit with the hash `"`r system2("git", "rev-parse --short HEAD", stdout = TRUE)`"`. To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).

@@

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of
R-based data analyses. To us Reproducibility means that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person later in time and the results will be identical to the ones intially reported [@herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability and a means to increase trustworthiness of empirical results [@epskamp2019rep].

Here, we combine four software tools, whose interplay can guarantee full computational reproducibility of data analyses and their reporting. There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018]. We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report. Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014]. Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity. They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice. However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure reproducibility of scientific analyses.

>>>>

affiliation:
  - id            : "1"
    institution   : "Center for Lifespan Psychology, Max Planck Institute for Human Development, Berlin, Germany"
  - id            : "2"
    institution   : "Humboldt-Universität zu Berlin"
  - id            : "3"
    institution   : "Max Planck UCL Centre for Computational Psychiatry and Ageing Research, Berlin, Germany & London, U.K." 
authornote: |
   We are grateful to Julia Delius for her helpful assistance in language and style editing. This paper is fully reproducible using the workflow described in this paper. All materials can be found on the accompanying GitHub repository: [https://github.com/aaronpeikert/reproducible-research/](https://github.com/aaronpeikert/reproducible-research/). To provide feedback or any other type of comment or questions regarding our workflow, please add an issue to the GitHub repository of our paper at [https://github.com/aaronpeikert/reproducible-research/issues](https://github.com/aaronpeikert/reproducible-research/issues).

@@

In this tutorial, we describe a workflow to ensure long-term and cross-platform reproducibility of data analyses in R [@baseR]. Reproducibility is the ability to draw identical conclusion from the same statistical analysis and the same data. In practice, reproducibility means to us that statistical results as well as their generating computational workflows are reported fully, and transparently, and remain sustainably available, such that a given workflow can be re-run by a different person, later in time, and that the results will be identical to the ones intially reported [@turingway2020; @herouxCompatibleReproducibilityTaxonomy2018; @claerboutElectronicDocumentsGive1992]. The need to ensure reproducibility directly follows from commonly accepted rules of good scientific practice [such as the guidelines of the German Research Foundation, @dfg2019]. Ensuring reproducibility is a prerequisite for replicability (the ability to reach a similar conclusion from the same analysis and new data), and a means to increase trustworthiness of empirical results [@epskamp2019rep].

Here, we combine four software tools, whose interplay can guarantee full computational reproducibility of data analyses and their reporting. There are various ideas on how to enhance reproducibility [@piccoloToolsTechniquesComputational2016], four of which we believe to be particularly important: literate programming: [@ruleTenSimpleRules2019], version control [@barbaHardRoadReproducibility2016], dependency management [@askrenUsingMakeReproducible2016], and containerization [@clyburne-sherinComputationalReproducibilityContainers2018]. We argue that only a workflow using all four concepts in unison can guarantee confidence in reproducing a scientific report [see @turingway2020 for similar arguments]. Various implementations of these concepts exist, but we consider the following four best suited for the R environment [@r2019]: for literate programming, R Markdown [@rmarkdown2019], for version control, Git [@chaconProGit2014], for dependency management, Make [@feldmanMakeProgramMaintaining1979], and for containerization, Docker [@merkelDockerLightweightLinux2014]. Each of these software solutions serves a valuable meta-scientific goal (reproducibility) and increases the researches' productivity. They all are highly flexible and powerful, so that mastering them fully, requires a considerable amount of practice. However, for our purposes, it is sufficient to master a valuable minimal subset of functions to ensure reproducibility of scientific analyses.
